---
title: "2018 Crescent Sail Yacht Club Regatta"
output: html_document
---

```{r setup, include=FALSE}
rm(list = ls()) # clear all environment variables
suppressPackageStartupMessages(library(leaflet))


knitr::opts_chunk$set(
	message = FALSE,
	warning = FALSE,
	include = FALSE,
	messages = FALSE
)
#library
library(tidyverse)
library(dplyr)
library(zoo)

exp_raw <- read.csv("../data/2018Aug25.csv", stringsAsFactors = FALSE)

#remove any columns where the data is entirely NA
allna <- which(apply(exp_raw, 2, function(x) all(is.na(x))))   #Find columns where all is not NA

exp <- exp_raw %>% select(-allna)

#Convert Time
exp$time <- as.POSIXct((exp$Utc) * (60*60*24), origin = "1899-12-30", tz = "America/Detroit")

#Select only needed columns
exp <- exp %>% select(time, Bsp, Awa, Aws, Twa, Tws, Twd, Hdg, PDOP, GeoSep, Lat, Lon, Cog, Sog, GPS.time, Twd.90, Mk.Lat, Mk.Lon, Stbd.lat, Stbd.lon)

#change names to lowercase for easier handling
names(exp) <- tolower(names(exp))

marks <- unique(exp[, c("mk.lon", "mk.lat")])
marks <- na.omit(marks)

#limit the data
exp <- exp[exp$time > as.POSIXct("2018-08-25 12:55:00 EDT"), ]


#Lets fill in some missing data
exp$sog <- na.approx(exp$sog, x=exp$time, na.rm = FALSE)
exp$lon <- na.approx(exp$lon, x=exp$time, na.rm = FALSE)
exp$lat <- na.approx(exp$lat, x=exp$time, na.rm = FALSE)
exp$twa <- na.approx(exp$twa, x=exp$time, na.rm = FALSE)
exp$tws <- na.approx(exp$tws, x=exp$time, na.rm = FALSE)


```




#Course Sailed
```{r message=FALSE, warning=FALSE}
#get states shape files
mi <- map_data("state") %>% filter(region=="michigan")



#plot the positions
m <- ggplot()+
  geom_polygon(data=mi, aes(long, lat, group=group)) + 
  coord_map("mercator", xlim=c(-82.9, -82.77), ylim=c(42.38, 42.475))



m + 
  geom_path(data=exp, aes(x=lon, y=lat, color=sog), size=2) + 
  geom_point(data=marks, aes(x=mk.lon, y=mk.lat), shape=17, size=3, color="orange") + 
  scale_color_viridis_c()

```

The tacks are supprisingly parallel. It seems the wind direction was relatively stable until later in the beat. 

```{r}

ggplot(exp) + 
  geom_path(aes(x=time, y=cog)) +
  scale_y_continuous(name="Bearing [Deg]")+
  ggtitle("Course Over Ground")

```

#Apparent Wind Angle
For some reason, the appaprent wind angle [AWA] and apparent wind speed [AWS] stopped logging around 1:50PM. Unfortunately, this means many of the values calculated from AWA and AWS can't be used past this time. This includes the true wind and polar targets. 

```{r}

ggplot(exp) + 
  geom_path(aes(x=time, y=awa)) +
  scale_y_continuous(name="Degree from bow")+
  ggtitle("Apparent Wind Angle") 

```

Apparent Wind Speed also stopped logging after some time. 
```{r}

ggplot(exp) + 
  geom_path(aes(x=time, y=aws)) +
  scale_y_continuous(name="KTS")+
  ggtitle("Apparent Wind Speed")

```


```{r}
ggplot(exp) + 
  geom_path(aes(x=time, y=twa)) +
  scale_y_continuous(name="Degree from bow") + 
  ggtitle("True Wind Angle")

```

```{r}
ggplot(exp) + 
  geom_path(aes(x=time, y=tws)) +
  scale_y_continuous(name="Degree from bow") + 
  ggtitle("True Wind Angle")

```

```{r}
ggplot(exp) + 
  geom_path(aes(x=time, y=sog)) +
  scale_y_continuous(name="KTS") + 
  ggtitle("Speed over Ground")

```






#Performance to Polar
```{r}
library(akima)

source("../src/optimalPerformanceFunctions.R")
load("../data/polarmodel.rda")

exp_opt <- exp[!is.na(exp$twa) & !is.na(exp$tws), ]

exp_opt$opt_v <- getOptV(btw = exp_opt$twa, vtw = exp_opt$tws, pol.model = pol.model)

exp_opt <- exp_opt[exp_opt$opt_v > 0, ]

exp_opt$opt_v_roll <- rollmean(exp_opt$opt_v, k = 10, fill = NA)

exp_opt$pol_perc <- (exp_opt$sog / exp_opt$opt_v_roll) * 100

ggplot(exp_opt) + 
  geom_path(aes(x=time, y=pol_perc, color=tws)) + 
  scale_color_viridis_c(name="True Wind Speed") +
  scale_x_datetime()+
  geom_vline(xintercept=as.POSIXct("2018-08-25 13:49:33"), color="red")+
  geom_hline(yintercept=100)

```

Only looking at the data where we have AWA. 
```{r}

good <- exp_opt[exp_opt$time < as.POSIXct("2018-08-25 13:49:33"), ]

poldata <- as.data.frame(pol.model$pol.matrix)
names(poldata) <- pol.model$tws
poldata <- gather(poldata, key = "tws", value="target")
poldata$twa <- pol.model$twa


ggplot(good[good$tws > 15.5 & good$tws < 16.5, ]) + 
  geom_point(aes(x=abs(twa), y=sog, color=pol_perc))+ 
  geom_path(data=poldata[poldata$tws==16, ], aes(x=twa, y=target, group=tws))+
  coord_polar() + 
  scale_x_continuous(limits=c(0,360))+
  ggtitle("TWS = 16 kts")

ggplot(good[good$tws > 19.5 & good$tws < 20.5, ]) + 
  geom_point(aes(x=abs(twa), y=sog, color=pol_perc), alpha=0.3)+ 
  geom_path(data=poldata[poldata$tws==16, ], aes(x=twa, y=target, group=tws))+
  coord_polar() + 
  scale_x_continuous(limits=c(0,360))+
  ggtitle("TWS = 20 kts")


```


